### 2017-03-14 latex/algorithmdesign/untitled-4.tex
finally made a 512-bit permutation

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Brief Article}
\author{The Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{First section}

Meet Bob and Alice. Bob and Alice are pen pals; they write letters to each other, and send them through the mail. Additionally, we will introduce Eve:  Eve has no respect for others, as well as a pathological need to try and do wrong; She is not opposed to lying, cheating, and stealing, and is very creative in this arts. She will go to great length to do these things if she thinks it will benefit her; Even when there is no real benefit to her, she will do such things whenever she can get away with it, simply because she can.

Unfortunately, Eve is the mail carrier. Fortunately, Bob and Alice are aware of Eve's nature, and they seek to find ways of communicating with each other without Eve interfering. In order to do so, they need to enumerate all of the ways that Eve could abuse their communications, and develop appropriate counter-measures.

\subsection{Lying, Cheating, and Stealing}

The first and most obvious concern is Eve opening and reading the messages that are exchanged between Bob and Alice.

Second, and slightly more devious, is that Eve could manipulate messages sent between the two.

Third, Eve actually does not have to wait for Bob to actually send a letter to Alice. She can simply write her own letter, put Bob's name on it, and deliver it to Alice.

Fourth, she could withhold or make copies of messages, and save them for delivery at some other arbitrary point in time.

Finally, Eve could simply choose not to deliver the messages at all.

\subsection{Solutions}

For example, Bob might send Alice a christmas card with money inside, and a note that says "I am sending you \$20; I hope you have a nice holiday". Upon opening this, Eve could simply take the money, and leave a new note in the card that say "Sorry, I can't afford gifts this year; I hope you have a nice holiday anyways. PS: Please send me some money". Then she could easily repeat this process.

\end{document}