### 2017-03-16 latex/algorithmdesign/untitled-4.tex

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[11pt]{article} % use larger type; default would be 10pt

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%%% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%%% PAGE DIMENSIONS
\usepackage{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} % for example, change the margins to 2 inches all round
% \geometry{landscape} % set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} % Activate to begin paragraphs with an empty line rather than an indent

%%% PACKAGES
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
% These packages are all incorporated in the memoir class to one degree or another...

%%% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%%% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%%% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

%%% END Article customizations

%%% The "real" document content comes below...

\title{Brief Article}
\author{The Author}
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\section{First section}

Meet Bob and Alice. Bob and Alice are pen pals; they write letters to each other, and send them through the mail. Additionally, we will introduce Eve:  Eve has no respect for others, as well as a pathological need to try and do wrong; She is not opposed to lying, cheating, and stealing, and is very creative in this arts. She will go to great length to do these things if she thinks it will benefit her; Even when there is no real benefit to her, she will do such things whenever she can get away with it, simply because she can.

Unfortunately, Eve is the mail carrier, and is responsible for delivering messages to/from the pair. Fortunately, Bob and Alice are aware of Eve's nature, and they seek to find ways of communicating with each other without Eve interfering. In order to do so, they need to enumerate all of the ways that Eve could abuse their communications, and develop appropriate counter-measures.

\subsection{Lying, Cheating, and Stealing}

The first and most obvious concern is Eve opening and reading the messages that are exchanged between Bob and Alice.

Second, and slightly more devious, is that Eve could manipulate messages sent between the two.

Third, Eve actually does not have to wait for Bob to actually send a letter to Alice. She can simply write her own letter, put Bob's name on it, and deliver it to Alice.

Fourth, she could withhold or make copies of messages, and save them for delivery at some other arbitrary point in time.

Finally, Eve could simply choose not to deliver the messages at all.

\subsection{Solutions}

In order to preclude Eve's ability to read their messages, Bob and Alice realize they need some way to obscure the information such that only the intended recipient may recover the actual information from the transported information. They need some kind of set of steps that they can apply to their message that will: 1: Produce output that is meaningless, and 2: The steps can be undone by the intended receipient to recover the actual message.

So clearly, they reason, they need some kind of technique that will allow the two of them be able to perform the steps, but nobody else (i.e. Eve) should be able to perform the steps. After some thought, they realize that if they can keep some part of the steps secret, and only the two of them know the secret steps, then clearly Eve will not be able perform the transformation and recover the message. The pair realize that Eve will go out of her way to attempt to figure out the secret steps, and that they will likely need to include many secrets; If they only use one secret, Eve could easily guess it.

% communicating the coin flip

\end{document}