### 2017-04-26 latex/keyexchange/secretkeybasedpublickeyencryption.tex
backing up work

\documentclass[preprint]{iacrtrans}
\usepackage[utf8]{inputenc}
\usepackage{titlesec}

% Select what to do with todonotes: 
% \usepackage[disable]{todonotes} % notes not showed
\usepackage[draft,color=orange!20!white,linecolor=orange,textwidth=3cm,colorinlistoftodos]{todonotes}   % notes showed
\setlength{\marginparwidth}{3cm}
\usepackage{graphicx}
\usepackage{soul}
\graphicspath{{images/}} % end dirs with `/'
% \usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{backgrounds}
\usetikzlibrary{arrows}
\usetikzlibrary{crypto.symbols}
\tikzset{shadows=no}        % Option: add shadows to XOR, ADD, etc.

\author{Ella Rose\inst{1}}
\institute{Paso Robles, CA \email{python_pride@protonmail.com}}
\title[\texttt EPQ Key Exchange]{\texttt EPQ Key Exchange}

\begin{document}

\maketitle

% use optional argument because the \LaTeX command breaks the PDF keywords
\keywords[Key Exchange]{Key Exchange}\todo{Keywords?}

\begin{abstract}
  We define an algorithm for performing a key exchange between two parties. The algorithm makes use of a small number of bignum multiplications and additions and is conceptually simple.\\ 
\end{abstract}

\todototoc
\listoftodos

\section{Introduction}
 The ability to exchange cryptographic keys over insecure channels is of vital importance to the function of the internet; Without algorithms for arranging secure communication online, much of every day life in the modern world would cease to be possible.

The introduction of Quantum Computing and post-quantum cryptography has lead to the development of lattice based cryptographic schemes that are believed to be secure even when faced with quantum parallelism. Simply put, RSA and DH will need replacement at some point, and lattices seem like a promising alternative.

Lattices are also very attractive because they offer homomorphic properties, which is another of the most important topics in cryptography at the moment. This property is important because it allows for the construction of simple public key encryption schemes and key exchange shemes from secret key ciphers. We present such a scheme in this paper, as well as a homomorphic cipher suitable for instantiating the scheme.

\section{Definitions}
We use $+$ to denote integer addition, $ab$ to denote integer multiplication, and $\mod p$ to denote the modulo operation. We use $log(x)$ to indicate the binary logarithm of $x$.

\section{Algorithm}
First, we shall define the asymmetric key exchange algorithm; Then, we shall define a symmetric secret key cipher suitable for instantiating the key exchange algorithm.

\subsection{Key Exchange Algorithm}
Let $p_1$ and $p_2$ be two encryptions of $0$ under the homomorphic secret key cipher. These two values constitute a public key. To share a secret with the owner of the public key, compute:
\begin{align}
    e \gets secret \\
    c \gets (p_1  r_1) + (p_2 r_2) + e
\end{align}

Where $r_1$ and $r_2$ are randomly generated values of appropriate size. Then the value $c$ is shared with the owner of the public key. The owner of the public key can recover $e$ from $c$ by applying the homomorphic decryption operation of the secret key cipher to c. Because $p_1$ is an encryption of $0$, $p_1 r_1 \equiv 0$, similar for $p_2$. So when the decryption operation is applied, the result is $0 + 0 + e \equiv e$.

\subsubsection{Additional Information for Key Generation}
If the size of the shared secret is to be 256 bits, then the ciphertexts produced by symmetric cipher must support the homomorphic addition of a plaintext 256-bit value.

Not all homomorphic constructions support the addition of plaintext values to ciphertext values; Lattices offer this required feature.

It is important that $p_1$ and $p_2$ are coprime to each other. Otherwise, some bits of the secret may be recovered by reducing a ciphertext modulo the greatest common divisor. Specifically:

\begin{align}
    e \gets secret\\
    c \gets (p_1 r_1) + (p_2 r_2) + e\\
    e_0 \gets c \mod gcd(p_1, p_2)
\end{align}

Where $log(e_0) \equiv log(gcd(p_1, p_2))$.

\subsection{Secret Key Cipher}
Creation of the $p_1$ and $p_2$ values requires a secure, additively homomorphic cipher. We define a simple lattice based cipher which utilizes effectively the same operations as the key exchange process. Let $s_1$ and $s_2$ be two large numbers, where $log(s_1) >= log(s_2) * 3$; These two numbers constitute the secret key for the cipher, which in turn acts as the private key for the key exchange presented previously.\\

To encrypt a ciphertext, compute:

\begin{align}
   c \gets (s_1 r_1) + (s_2 r_2) + m
\end{align}

Where the $r$ values are of appropriate size relative to $s_1$ and $s_2$ \todo{what sizes are appropriate}
While decryption is computed as:

\begin{align}
    m \gets (c \mod s_1) \mod s_2
\end{align}

\subsubsection{Additional Information for Key Generation}
$s_1$ and $s_2$ should be coprime, and $log(s_1) >= log(s_2) * 3$

\subsection{Note}
While the decryption circuit works via two modulo operations of the base lattice points, this operation does not work for the public key/key exchange portion of the algorithm - The reason is because in the secret key cipher, the two different lattice points are of largely different size; In the key exchange algorithm, both points are roughly the same size, and so the double modulo operation does not function identically.

\subsection{Public Key Randomization}
Because the public key consists of malleable homomorphic ciphertexts, it can be randomized without complicating any of the rest of the procedures. The presented scheme can utilize encrypted public keys, and we give two potential use cases for the idea.

First, as an additional layer of security, users who wish to send a message to the owner of the public key can first randomize the public key that they are using. Assuming that the randomized $p_1$ and $p_2$ are in effect unrelated to the original $p_1$ and $p_2$, then observers of the ciphertext do not even have the base points that the ciphertext was created with.

Secondly, this feature allows for the disassociation of the owner of the public key from the possessor of the public key. For example, assume we have three parties, Alice, Bob, and Charlie. Charlie does not want Alice to communicate with Bob. If Charlie searches Alices computer and he finds Bobs public key, then he will know that Alice has been or intended to communicate with Bob. To counter this, Alice can randomize Bobs public key and store only the randomized key. The randomized key will be functionally equivalent to the public key associated with Bob, with the exception that nobody may associate it with Bob.

\section{Conclusion}
 We define a simple lattice based key exchange algorithm. The algorithm utilizes 4 bignum multiplications and 4 bignum additions to generate a keypair, 2 bignum multiplications and 2 bignum additions per key sent, and 2 modulo operations on bignums to receive a key. Public keys occupy < 1200 bits using current parameter cohices, while private keys require about 600 bits.
\end{document}